Portfolios of Subgraph Isomorphism Algorithms

### 2015-12-14 paper/paper.tex
General editing

\usepackage{times}
\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{cleveref}                  % no need to type Figure etc
\usepackage{xcolor,colortbl}
\usepackage{tabularx}

@@

\begin{abstract}
Subgraph isomorphism is a computationally challenging problem with lots of important practical
applications, for example in computer vision, biochemistry, and model checking. There are a number
of state-of-the-art algorithms for solving it, each of which has its own performance
characteristics. As with many other hard problems, the single best choice of algorithm overall is
rarely the best algorithm for each instance. We develop a portfolio approach that leverages novel
features to characterise subgraph isomorphism problems to dynamically decide which algorithm to use
on a per-instance basis. We demonstrate significant performance improvements on a large set of hard
benchmark problems. In addition, we show how algorithm selection models can be leveraged to gain new
insights into what affects the performance of an algorithm.
\end{abstract}

@@

The subgraph isomorphism problem is to find an injective mapping from vertices of a small
\emph{pattern} graph to vertices of a large \emph{target} graph which preserves adjacency. This
\NP-complete problem has lots of important practical applications, for example in computer vision
\cite{cviu11,pr15}, biochemistry \cite{Giugno:2013}, and model checking \cite{Sevegnani:2015}. There
exist various exact algorithms, which have been compared on a large benchmark in
\cite{McCreesh:2015}. These experiments indicated that the single best algorithm depends on the CPU
time limit considered: for very small (resp. larger) CPU time limits, VF2 \cite{Cordella:2004}
(resp. the Glasgow algorithm \cite{McCreesh:2015}) has better success rates. Furthermore, they also
showed us that on an instance by instance basis, other algorithms may become better.

@@

A \emph{graph} $G=(N,E)$ consists of a \emph{node set} $N$ and an \emph{edge set} $E \subseteq N \times N$, where an edge $(u,u')$ is a couple of nodes. The number of neighbors of a node $u$ is the degree of $u$, $d^\circ(u)=\#\{ (u,u')\in E\}$. In this paper, we implicitely consider non directed graphs, such that $(u,u')\in E\Leftrightarrow (u',u)\in E$. The extension to directed graphs is rather straightforward, and all algorithms compared in this paper can handle directed graphs as well.

A \emph{subgraph isomorphism problem} between a pattern graph $G_p=(N_p,E_p)$ and a target graph $G_t=(N_t,E_t)$ consists in deciding whether $G_p$ is isomorphic to some subgraph of $G_t$. More precisely, the goal is to find an injective matching $f: N_p\rightarrow N_t$, that associates a different target node to each pattern node, and that preserves pattern edges, i.e., $\forall (u,u') \in E_p, (f(u),f(u')) \in E_t$.

Note that the subgraph is not necessarily induced so that two pattern nodes that are not linked by an edge may be matched to two target nodes which are linked by an edge. This problem is also called subgraph monomorphism or subgraph matching in the literature. 
%{\em Induced subgraph isomorphism} is a particular case of subgraph isomorphism such that the subgraph isomorphism function must also preserve target edges between matched nodes, i.e., 
%\[\forall (u,u') \in N_p^2, (u,u')\in E_p \Leftrightarrow(f(u),f(u')) \in E_t\]

In the following, we assume $G_p=(N_p,E_p)$ and $G_t=(N_t,E_t)$ to be the underlying instance of subgraph isomorphism problem. 
We also define $n_p = \# N_p$, $n_t = \# N_t$,  $e_p=\# E_p$, $e_t=\# E_t$, and $d_p$ and $d_t$ the maximal degrees of the graphs $G_p$ and $G_t$.

@@

Subgraph isomorphism problems may be solved by a systematic exploration of the search space composed of all possible injective matchings from $N_p$ to $N_t$: starting from an empty matching, one incrementally extends a partial matching by matching a non matched pattern node to a non matched target node until either some edges are not matched by the current matching (the search must backtrack to a previous choice point and go on with another extension) or all pattern nodes have been matched (a solution has been found). To reduce the search space, this exhaustive exploration is combined with filtering techniques that aim at removing candidate couples of non matched pattern-target nodes $(u,v)\in N_p\times N_t$. Different filterings may be considered; some are stronger than others (they remove more candidate couples), but also have higher time complexities.

@@

Simple filterings basically propagate difference constraints (which ensure that the matching is injective) and edge constraints (which ensure that the matching preserves pattern edges): each time a pattern node $u\in N_p$ is matched with a target node $v\in N_t$, one removes every candidate couple $(u',v')\in N_p\times N_t$ such that either $v'=v$ (difference constraint) or $(u,u')$ is a pattern edge whereas $(v,v')$ is not a target edge (edge constraint). This simple filtering (called {\em Forward-Checking}) is very fast to achieve: in ${\cal O}(n_p)$ for difference constraints, and in ${\cal O}(d_p\cdot n_t)$ for edge constraints. It is used, for example, in McGregor's algorithm \cite{mcgregor79} and in VF2 \cite{Cordella:2004}.

R\'egin has introduced in \cite{regin} a stronger filtering for difference constraints, which ensures that all pattern nodes can be matched with different target nodes, all together. This filtering (called {\em AllDifferent Generalized Arc Consistency}) removes more candidate couples than when each difference contraint is propagated separately which Forward-Checking. However, it is also more time consuming as it is done in ${\cal O}(n_p^2\cdot n_t^2)$.

Various filterings have been introduced for edge constraints. Ullman \cite{ullman} has introduced a filtering which basically ensures that for each pattern edge $(u,u')\in E_p$ and each candidate couple $(u,v)\in N_p\times N_t$, there exists a candidate couple $(u',v')\in N_p\times N_t$ such that $(v,v')$ is a target edge. Candidate couples $(u,v)$ that do not satisfy this property are iteratively removed until a fixed point is reached. This filtering (called {\em Arc Consistency}) removes more candidate couples than Forward-Checking, but it is also more time consuming as it is done in ${\cal O}(e_p\cdot n_t^2)$ when using the algorithm AC4  \cite{MH86}.

A stronger filtering may be obtained by propagating edge constraints in a more global way, as proposed in \cite{LV02}. The idea is to check for each candidate couple $(u,v)\in N_p\times N_t$ that the number of pattern nodes adjacent to $u$ is smaller than or equal to the number of target nodes that are both adjacent to $v$ and that may be matched with nodes adjacent to $u$. This is done in ${\cal O}(n_p^2\cdot n_t^2)$. This idea has been generalized in \cite{Solnon:2010} where, for each candidate couple $(u,v)\in N_p\times N_t$, a redundant Local AllDifferent (LAD) constraint ensures that each neighbour of $u$ may be matched with a different neighbour of $v$. This is done in ${\cal O}(n_p\cdot n_t\cdot d_p^2\cdot d_t^2)$.

@@

Some filterings exploit invariant properties, {\em i.e.}, properties associated with nodes such that nodes may be matched only if they have compatible properties. A classical property is the degree: a pattern node $u\in N_p$ may be matched with a target node $v\in N_t$ only if $d^\circ(u)\leq d^\circ(v)$. This property is usually used at the beginning of the search to reduce the set of candidate couples to $\{(u,v)\in N_p\times N_t\;|\;d^\circ(u)\leq d^\circ(v)\}$.
Other examples of invariant properties are the number of cycles of length $k$ passing through the node, and the number of cliques of size $k$ containing the node, which must be smaller for a pattern node than for its matched target node.
Invariant properties may also be associated with couples of nodes. For example, the number of paths of length $k$ between two pattern nodes is smaller than or equal to the number of paths of length $k$ between the target nodes they may be matched with.

These invariant properties are used, for example:
\begin{itemize}
\item  In \cite{battiti-mascia07}, to remove candidate couples $(u,v)\in N_p\times N_t$ such that the number of paths starting from  pattern node $u$ is greater than the number of paths starting from  target node $v$~;
\item In \cite{Audemard:2014} to generalize the LAD constraint proposed in \cite{Solnon:2010} so that it ensures that a subset of pattern nodes can be matched with all different compatible target nodes, where compatibility is defined with respect to invariant properties~;
\item In \cite{McCreesh:2015} to filter the set of candidate couples before starting the search.
\end{itemize}
In \cite{Audemard:2014} , the length of paths is not limited, and it is iteratively incremented until no more couple is removed. In \cite{battiti-mascia07} and \cite{McCreesh:2015},  algorithms are parameterized by the maximum path length considered when counting paths: Larger values for this parameter remove more candidate couples, but are also more time consuming. Experiments reported in \cite{battiti-mascia07} show us that the best setting depends on the instance considered, and that a portfolio running several randomized 
versions in time-sharing decreases the total CPU-time needed to find a solution for feasible instances. In experiments reported in \cite{McCreesh:2015}, this parameter has been set to 3, as this setting seemed to be a reasonable compromise.

@@

The per-instance algorithm selection problem~\cite{rice_algorithm_1976} is to select from an
algorithm portfolio~\cite{huberman_economics_1997,gomes_algorithm_2001} the one expected to perform
best on a given problem instance. Algorithm selection systems usually build machine learning models
of the algorithms or the portfolio they are contained in to forecast which algorithm to use in a
particular context. Using the model predictions, one or more algorithms from the portfolio are
selected to be run sequentially or in parallel.

@@

\begin{itemize}
\item Class 1 contains randomly generated scale-free graphs \cite{constraints10}.
\item Classes 2 and 3 contain various kinds of graphs coming from \cite{LV02}: class 2 contains small
    instances generated from the 50 first graphs, whereas class 3 contains larger
    instances where pattern graphs are coming from class 2, and target graphs are coming from the
    next 50 graphs.
\item Classes 4 to 8 contain randomly generated graphs coming from
    \cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for classes 3 and 4, regular meshes
    for classes 5 and 6, and random graphs with uniform edge probabilities for class 7.
\item Classes 9 and 10 contain instances coming from segmented images \cite{pr15,cviu11}.
\item Class 11 contains instances coming from meshes modeling 3D objects \cite{cviu11}.
\item Class 12 contains random graph instances belonging to the phase transition.
\end{itemize}

Note that Classes 3 and 12 were not considered in the previous experimental
study reported in \cite{McCreesh:2015}.  This set of instances is also much
larger than that of Battiti and Mascia~\cite{battiti-mascia07}, who were the
first to propose algorithm portfolios for subgraph isomorphism problems.
Battiti and Mascia only considered a pure parallel portfolio consisting of two
solvers, and without a selection mechanism. Their problem set consisted
entirely of easy satisfiable instances, which is where this style of approach
has most success.

@@

We have selected a set of algorithms with complementary performance:
\begin{itemize}
\item VF2 \cite{Cordella:2004}, which performs a weak filtering so that it is especially fast on trivially satisfiable instances;
\item LAD \cite{Solnon:2010}, which combines two strong but expensive filtering techniques (AllDifferent Generalized Arc Consistency and LAD);
\item Glasgow \cite{McCreesh:2015}, which does expensive preprocessing based on path length
    invariant properties to generate additional constraints, followed by weaker filtering
    (forward-checking, and a heuristic AllDifferent propagator which can miss deletions) and
    conflict-directed backjumping during search.
\end{itemize}

\noindent The Glasgow algorithm has a parameter, which controls the lengths of paths used when
reasoning about non-adjacent vertices.  In experiments reported in \cite{McCreesh:2015}, the choice
of paths of length 3 was used as a reasonable compromise---longer paths lead to prohibitively
expensive preprocessing on larger, denser instances. This is often not the best choice on an
instance by instance basis: sometimes path-based reasoning gives no benefit at all, sometimes
considering only paths of length 2 suffices, occasionally paths of length 4 are helpful, and even
looking at paths of length 3 is relatively expensive on some graphs.

We also introduce two new variants of LAD. The first variant is a weaker LAD filtering (called incompleteLAD) which is applied once, without performing a backtracking search, and very quickly detects inconsistencies on many instances: For each pattern node $u$, we check that there exists at least one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor $v'$ of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$. IncompleteLAD is an incomplete algorithm as it can only detect inconsistency of some instances: when it does not detect inconsistency, the instance may either be consistent or not. As a counterpart of this incompleteness, it is very fast: its time complexity is ${\cal O}(n_p(n_t+e_t))$.

The second variant of LAD is called pathLAD and it combines the LAD constraints introduced in \cite{Solnon:2010} with the exploitation of path length properties, as proposed in \cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of paths of length 2 between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$, and to add the constraint that the label of a pattern node (resp. edge) must be smaller than or equal to the label of its associated target node (resp. edge).

@@

Figure \ref{expTimeGraph} displays the evolution of the number of instances solved with respect to CPU time. It shows us that the best solver depends on the time limit
considered. IncompleteLAD is able to solve easy inconsistent instances very quickly, in a few milliseconds. Hence, for time limits lower than 5 ms, the best solver is incompleteLAD. However, it is not able to solve harder inconsistent instances, neither can it solve consistent instances.

PathLAD and Glasgow(1) outperform incompleteLAD when considering longer time limits: PathLAD is the best solver for time limits larger than 5ms and lower than 40 ms, and Glasgow(1) is the best solver for time limits larger than 40ms and lower than 3000 ms.

Finally, for CPU time limits larger than 3000 ms, the best solver becomes Glasgow(2).
As we increase the CPU time limit, variants of Glasgow with longer paths (Glasgow(3) and Glasgow(4)) become better. This is what we expect, as more reasoning is expensive, but potentially increases the reduction of the search space. In particular, the relative performance of Glasgow variants is monotone in the time it takes to solve the instance – Glasgow(1) is best at first, but gets worse as the time to solve the instance increases until Glasgow(2) is better. After that point, Glasgow(1) is never better than Glasgow(2) again. The same is true for path lengths 3 and 4.

@@

Furthermore, the Virtual Best Solver (VBS), which considers the best algorithm for each instance
separately, obtains much better results, showing us that the algorithms have complementary
performance. The difference between VBS and single best is more particularly important for rather
small CPU time limits (smaller than 1000 ms). In many applications, it is important to have the
fastest possible algorithm. For example, in pattern recognition applications such as those reported
in \cite{pr15,cviu11}, and in chemical applications such as \cite{Giugno:2013}, we often have to
solve subgraph isomorphism problems for a very large number of graphs (in order to find a pattern
image or molecule in a large database of target images or compounds, for example), so that having an
algorithm that is able to solve an instance in 100 ms instead of 1000 ms makes a big difference.
Therefore, it is important to select the best algorithm for each instance, even if the instance is
an easy one.

Table \ref{expClass} shows us that we cannot reasonably split things up based upon which class instances are coming from, unlike SAT. For all classes, there are always at least 2 algorithms which are the best for at least one instance of the class. In particular, for classes 2 and 3, each algorithm is the best for at least one instance (except Glasgow 4 for class 3).

@@

Our approach is composed of 3 steps described below. First, we run two presolvers in a static way to
quickly solve easy instances; second, we extract features from instances which are not solved by the
first step; and finally, we select an algorithm and run it.

@@

Experimental results reported in Section \ref{expComp} have shown us that incompleteLAD is very fast
(7 ms on average) and is able to very quickly solve 1919 instances. Therefore, we first run
incompleteLAD: if the instance is detected to be unsatisfiable, then we do not process it further.
Otherwise, we record as features the number of candidate couples removed by incompleteLAD in absolute
terms, as a percentage, and the minimum and maximum on a per-variable basis.

VF2 is also able to solve many easy instances very quickly: among the 3806 instances which are not
solved by incompleteLAD, 1470 are solved by VF2 in less than 50 milliseconds. Therefore, after
running incompleteLAD, we run the VF2 solver for 50 ms. This solves easy instances without the
overhead of running algorithm selection and avoids potentially making incorrect solver choices.
After running incompleteLAD and VF2 for 50 ms, we are left with 2336 hard instances that we consider
for algorithm selection.

@@

LLAMA's pairwise regression approach with random forest regression gave the best
performance. The idea is very similar to the pairwise classification used
in~\cite{xu_satzilla_2008}. For each pair of algorithms in our portfolio, we
train a model that predicts the performance difference between them. That is, if
the first algorithm is better than the second, the difference is positive,
otherwise negative. The algorithm with the highest cumulative performance
difference, i.e.\ the most positive difference over all other algorithms, is
chosen to be run.

@@

Table~\ref{tab:res} shows the performance of our algorithm selection approach, compared to two
baselines. The virtual best solver is the oracle predictor that, for each instance, chooses the best
solver from our portfolio. This is the upper bound of what an algorithm selection approach can
achieve. The single best solver is the one solver from the portfolio that has the overall best
performance across the entire set of instances, at the CPU time limit of $10^8$ ms, {\em i.e.}, Glasgow(2). We consider it a lower bound on performance.

\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \toprule
model & mean MCP & solved instances & mean performance\\
  \midrule
virtual best & 0 & 2219 & 5822809\\
single best & 1965158 & 2172 & 7787967\\
LLAMA & 610835 & 2206 & 6435301\\
   \bottomrule
\end{tabular}
\vspace{1ex}
\caption{Algorithm selection performance on the reduced set of 2336
instances. MCP is the misclassification penalty; that is, the additional time
required to solve a set of instances because of choosing solvers that perform
worse than the best.}\label{tab:res} \end{table}

We are able to solve 34 more instances than the single best solver within the
timeout, with only an additional 13 to the virtual best. In terms of average
performance, we are able to close almost 70\% of the gap between the single best
and the virtual best solver.

>>>>

\usepackage{times}
\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{cleveref}                  % no need to type Figure etc
\usepackage{xcolor,colortbl}
\usepackage{tabularx}
\usepackage{siunitx}

@@

\begin{abstract}
Subgraph isomorphism is a computationally challenging problem with important practical
applications, for example in computer vision, biochemistry, and model checking. There are a number
of state-of-the-art algorithms for solving the problem, each of which has its own performance
characteristics. As with many other hard problems, the single best choice of algorithm overall is
rarely the best algorithm on an instance-by-instance. We develop a portfolio approach which leverages
novel features to characterise subgraph isomorphism problems to dynamically decide which algorithm
to use on a per-instance basis. We demonstrate significant performance improvements on a large set
of hard benchmark problems. In addition, we show how algorithm selection models can be leveraged to
gain new insights into what affects the performance of an algorithm.
\end{abstract}

@@

The subgraph isomorphism problem is to find an adjacency-preserving injective mapping from vertices
of a small \emph{pattern} graph to vertices of a large \emph{target} graph. This \NP-complete
problem has many important practical applications, for example in computer vision
\cite{cviu11,pr15}, biochemistry \cite{Giugno:2013}, and model checking \cite{Sevegnani:2015}. There
exist various exact algorithms, which have been compared on a large suite of problem instances by
McCreesh and Prosser \cite{McCreesh:2015}. These experiments indicated that the single best
algorithm depends on the CPU time limit considered: for very small CPU time limits, VF2
\cite{Cordella:2004} is the best choice, whereas the Glasgow algorithm \cite{McCreesh:2015} has
better success rates for larger time limits.  They also showed that on an instance by instance
basis, other algorithms are often better.

@@

A \emph{graph} $G=(N,E)$ consists of a \emph{node set} $N$ and an \emph{edge set} $E \subseteq N
\times N$, where an edge $(u,u')$ is a pair of nodes. The number of neighbors of a node $u$ is
called the degree of $u$, $d^\circ(u)=\#\{ (u,u')\in E\}$. In this paper, we implicitly consider
non-directed graphs, such that $(u,u')\in E\Leftrightarrow (u',u)\in E$. The extension to directed
graphs is rather straightforward, and all algorithms compared in this paper can handle directed
graphs as well.

Given a pattern graph $G_p=(N_p,E_p)$ and a target graph $G_t=(N_t,E_t)$, the \emph{subgraph
isomorphism problem} consists of deciding whether $G_p$ is isomorphic to some subgraph of $G_t$.
More precisely, the goal is to find an injective matching $f: N_p\rightarrow N_t$, that associates a
different target node to each pattern node, and that preserves pattern edges, i.e.\ $\forall (u,u')
\in E_p, (f(u),f(u')) \in E_t$.

Note that the subgraph is not necessarily induced, so that two pattern nodes that are not linked by
an edge may be mapped to two target nodes which are linked by an edge. This problem is also called
\emph{subgraph monomorphism} or \emph{subgraph matching} in the literature.

In the following, we assume $G_p=(N_p,E_p)$ and $G_t=(N_t,E_t)$ to be the underlying instance of a
subgraph isomorphism problem.  We also define $n_p = \# N_p$, $n_t = \# N_t$,  $e_p=\# E_p$, $e_t=\#
E_t$, and $d_p$ and $d_t$ to be the maximum degrees of the graphs $G_p$ and $G_t$.

@@

Subgraph isomorphism problems may be solved by a systematic exploration of the search space composed
of all possible injective matchings from $N_p$ to $N_t$: starting from an empty matching, one
incrementally extends a partial matching by matching a non-matched pattern node to a non-matched
target node until either some edges are not matched by the current matching (so the search must
backtrack to a previous choice point and go on with another extension), or all pattern nodes have
been matched (a solution has been found). To reduce the search space, this exhaustive exploration is
combined with filtering techniques that aim at removing candidate pairs of non-matched
pattern-target nodes $(u,v)\in N_p\times N_t$. Different filtering techniques may be considered;
some are stronger than others (they remove more candidate pairs), but also have higher time
complexities.

@@

The simplest form of filtering is simply to propagate difference constraints (which ensure that the
matching is injective) and edge constraints (which ensure that the matching preserves pattern
edges): each time a pattern node $u\in N_p$ is matched with a target node $v\in N_t$, one removes
every candidate pair $(u',v')\in N_p\times N_t$ such that either $v'=v$ (difference constraint) or
$(u,u')$ is a pattern edge but $(v,v')$ is not a target edge (edge constraint). This simple
filtering (called \emph{Forward-Checking}) is very fast to achieve: in ${\cal O}(n_p)$ for
difference constraints, and in ${\cal O}(d_p\cdot n_t)$ for edge constraints. It is used, for
example, in McGregor's algorithm \cite{mcgregor79} and in VF2 \cite{Cordella:2004}.

R\'egin \cite{regin} introduced a stronger filtering for difference constraints, which ensures that
all pattern nodes can be matched with different target nodes, all together. This filtering (called
\emph{AllDifferent Generalized Arc Consistency}) removes more candidate pairs than when each
difference constraint is propagated separately which Forward-Checking. However, it is also more time
consuming as it is done in ${\cal O}(n_p^2\cdot n_t^2)$ time.

Various filtering techniques have been tried for edge constraints. Ullman \cite{ullman} introduced a
filtering which ensures that for each pattern edge $(u,u')\in E_p$ and each candidate pair $(u,v)\in
N_p\times N_t$, there exists a candidate couple $(u',v')\in N_p\times N_t$ such that $(v,v')$ is a
target edge. Candidate couples $(u,v)$ that do not satisfy this property are iteratively removed
until a fixed point is reached. This filtering (called \emph{Arc Consistency}) removes more
candidate pairs than Forward-Checking, but it is also more time consuming as it is done in ${\cal
O}(e_p\cdot n_t^2)$ when using the algorithm AC4  \cite{MH86}.

Stronger filtering may be obtained by propagating edge constraints in a more global way, as proposed
by Larrosa and Valiente \cite{LV02}. The idea is to check for each candidate pair $(u,v)\in
N_p\times N_t$ that the number of pattern nodes adjacent to $u$ is smaller than or equal to the
number of target nodes that are both adjacent to $v$ and that may be matched with nodes adjacent to
$u$. This is done in ${\cal O}(n_p^2\cdot n_t^2)$. This idea was generalised by Solnon
\cite{Solnon:2010} where, for each candidate pair $(u,v)\in N_p\times N_t$, a redundant Local
AllDifferent (LAD) constraint ensures that each neighbour of $u$ may be matched with a different
neighbour of $v$. This is done in ${\cal O}(n_p\cdot n_t\cdot d_p^2\cdot d_t^2)$.

@@

Some filtering techniques exploit invariant properties, i.e.\ properties associated with nodes such
that nodes may be matched only if they have compatible properties. A classical property is the
degree: a pattern node $u\in N_p$ may be matched with a target node $v\in N_t$ only if
$d^\circ(u)\leq d^\circ(v)$. This property is usually used at the beginning of the search to reduce
the set of candidate pairs to $\{(u,v)\in N_p\times N_t\;|\;d^\circ(u)\leq d^\circ(v)\}$.  Other
examples of invariant properties are the number of cycles of length $k$ passing through the node,
and the number of cliques of size $k$ containing the node, which must be smaller for a pattern node
than for its matched target node.  Invariant properties may also be associated with pairs of nodes.
For example, the number of paths of length $k$ between two pattern nodes is smaller than or equal to
the number of paths of length $k$ between the target nodes with which they may be matched.

These invariant properties are used, for example:

\begin{itemize}
\item  By Battiti and Mascia \cite{battiti-mascia07}, to remove candidate pairs $(u,v)\in N_p\times
    N_t$ such that the number of paths starting from  pattern node $u$ is greater than the number of
    paths starting from  target node $v$;
\item By Audemard et al.\ \cite{Audemard:2014} to generalise the LAD constraint proposed by Solnon
    \cite{Solnon:2010} so that it ensures that a subset of pattern nodes can be matched with all
    different compatible target nodes, where compatibility is defined with respect to invariant
    properties;
\item by McCreesh and Prosser \cite{McCreesh:2015} to filter the set of candidate couples before
    starting the search, and to generate additional implied adjacency-like constraints which are
    processed during search.
\end{itemize}

\noindent Audemard et al.\ \cite{Audemard:2014} do not limit the length of paths considered, and
iteratively increment the length until no more pairs are removed. Battiti and Mascia
\cite{battiti-mascia07}, and McCreesh and Prosser \cite{McCreesh:2015} parameterise their algorithms
by the maximum path length considered when counting paths: larger values for this parameter remove
more candidate pairs, but are also more time consuming. Battiti and Mascia's experiments show that
the best setting depends on the instance considered, and that a portfolio running several randomised
versions in time-sharing decreases the total CPU time needed to find a solution for feasible
instances. McCreesh and Prosser simply set the parameter to 3, as this setting seemed to be a
reasonable compromise.

@@

The per-instance algorithm selection problem~\cite{rice_algorithm_1976} is to select from an
algorithm portfolio~\cite{huberman_economics_1997,gomes_algorithm_2001} the one expected to perform
best on a given problem instance. Algorithm selection systems usually build machine learning models
of the algorithms or the in which portfolio they are contained to forecast which algorithm to use in a
particular context. Using the model predictions, one or more algorithms from the portfolio are
selected to be run sequentially or in parallel.

@@

\begin{itemize}
\item Class 1 contains randomly generated scale-free graphs \cite{constraints10}.
\item Classes 2 and 3 contain various kinds of graphs selected by Larrosa and Valiente \cite{LV02}:
    class 2 contains small instances generated from the 50 first graphs, and class 3 contains
    larger instances where pattern graphs are coming from class 2, and target graphs are coming from
    the next 50 graphs.
\item Classes 4 to 8 contain randomly generated graphs coming from a database of graphs commonly
    used for benchmarking subgraph isomorphism algorithms
    \cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for classes 3 and 4, regular meshes
    for classes 5 and 6, and random graphs with uniform edge probabilities for class 7. All of these
    instances are satisfiable.
\item Classes 9 and 10 contain instances coming from segmented images \cite{pr15,cviu11}.
\item Class 11 contains instances coming from meshes modeling 3D objects \cite{cviu11}.
\item Class 12 contains random graph instances chosen to be close to the satisfiable-unsatisfiable
    phase transition---these instances are expected to be particularly challenging, despite their
    small size.
\end{itemize}

Note that Classes 3 and 12 were not considered in the previous experimental study by McCreesh and
Prosser \cite{McCreesh:2015}.  This set of instances is also much larger than that of Battiti and
Mascia~\cite{battiti-mascia07}, who were the first to propose algorithm portfolios for subgraph
isomorphism problems.  Battiti and Mascia only considered a pure parallel portfolio consisting of
two solvers, and without a selection mechanism. Their problem set consisted entirely of easy
satisfiable instances, which is where this style of approach has most success.

@@

We have selected a set of algorithms with complementary performance:
\begin{itemize}
\item VF2 \cite{Cordella:2004}, which performs a weak filtering that it is especially fast on
    trivially satisfiable instances;
\item LAD \cite{Solnon:2010}, which combines two strong but expensive filtering techniques
    (AllDifferent Generalized Arc Consistency and LAD);
\item Glasgow \cite{McCreesh:2015}, which does expensive preprocessing based on path length
    invariant properties to generate additional constraints, followed by weaker filtering
    (forward-checking, and a heuristic AllDifferent propagator which can miss deletions) and
    conflict-directed backjumping during search.
\end{itemize}

\noindent The Glasgow algorithm has a parameter, which controls the lengths of paths used when
reasoning about non-adjacent vertices.  In experiments reported by McCreesh and Prosser
\cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable compromise---longer
paths lead to prohibitively expensive preprocessing on larger, denser instances. This is often not
the best choice on an instance by instance basis: sometimes path-based reasoning gives no benefit at
all, sometimes considering only paths of length 2 suffices, occasionally paths of length 4 are
helpful, and even looking at paths of length 3 is relatively expensive on some graphs.

We also introduce two new variants of LAD. The first variant is a weaker LAD filtering (called
incompleteLAD) which is applied once, without performing a backtracking search, and very quickly
detects inconsistencies on many instances: for each pattern node $u$, we check that there exists at
least one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor
$v'$ of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$.
IncompleteLAD is an incomplete algorithm as it can only detect inconsistency of some instances: when
it does not detect inconsistency, the instance may either be satisfiable or not. As a counterpart of
this incompleteness, it is very fast: its time complexity is ${\cal O}(n_p(n_t+e_t))$.

The second variant of LAD is called pathLAD, which combines the LAD constraints introduced by Solnon
\cite{Solnon:2010} with the exploitation of path length properties proposed by Audemard et
al.\ \cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of paths of length 2
between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$,
and to add the constraint that the label of a pattern node (resp.\ edge) must be smaller than or
equal to the label of its associated target node (resp.\ edge).

@@

Figure \ref{expTimeGraph} displays the evolution of the number of instances solved with respect to
CPU time. It shows us that the best solver depends on the time limit considered. IncompleteLAD is
able to solve easy unsastisfiable instances very quickly, in a few milliseconds. Hence, for time
limits lower than \SI{5}{\ms}, the best solver is incompleteLAD. However, it is not able to solve harder
unsatisfiable instances, nor can it solve satsifiable instances.

PathLAD and Glasgow(1) outperform incompleteLAD when considering longer time limits: PathLAD is the
best solver for time limits larger than \SI{5}{\ms} and lower than \SI{40}{\ms}, and Glasgow(1) is the best solver
for time limits larger than \SI{40}{\ms} and lower than \SI{3000}{\ms}.

For CPU time limits larger than \SI{3000}{\ms}, the best solver becomes Glasgow(2).  As we
increase the CPU time limit, variants of Glasgow with longer paths (Glasgow(3) and Glasgow(4))
become better. This is what we expect, as more reasoning is expensive, but potentially increases the
reduction of the search space. In particular, the relative performance of Glasgow variants is
monotone in the time it takes to solve the instance---Glasgow(1) is best at first, but gets worse as
the time to solve the instance increases until Glasgow(2) is better. After that point, Glasgow(1) is
never better than Glasgow(2) again. The same is true for path lengths 3 and 4.

@@

Furthermore, the Virtual Best Solver (VBS), which considers the best algorithm for each instance
separately, obtains much better results, showing us that the algorithms have complementary
performance. The difference between VBS and single best is more particularly important for rather
small CPU time limits (smaller than \SI{1000}{\ms}). In many applications, it is important to have the
fastest possible algorithm. For example, in pattern recognition \cite{pr15,cviu11} and chemical
\cite{Giugno:2013} applications, we often have to solve the subgraph isomorphism problem repeatedly
for a very large number of graphs (in order to find a pattern image or molecule in a large database
of target images or compounds, for example), so having an algorithm that is able to solve an
instance in \SI{100}{\ms} instead of \SI{1000}{\ms} makes a big difference.  Therefore, it is important to select
the best algorithm for each instance, even if the instance is an easy one.

Table~\ref{expClass} shows us that we cannot simply select algorithms based upon which class from
which instances are coming (unlike in SAT, where this kind of approach can be reasonable). For all
classes, there are always at least two algorithms which are the best for at least one instance of
the class. In particular, for classes 2 and 3, each algorithm is the best for at least one instance
(except Glasgow(4) for class 3).

@@

Our approach is composed of three steps. First, we run two presolvers in a static way to quickly
solve easy instances; second, we extract features from instances which are not solved by the first
step; and third, we select an algorithm and run it.

@@

Experimental results reported in Section~\ref{expComp} have shown us that incompleteLAD is very fast
(\SI{7}{\ms} on average) and is able to solve 1919 instances very quickly. Therefore, we first run
incompleteLAD: if the instance is detected to be unsatisfiable, then we do not process it further.
Otherwise, we record as features the number of candidate couples removed by incompleteLAD in absolute
terms, as a percentage, and the minimum and maximum on a per-variable basis.

VF2 is also able to solve many easy instances very quickly: among the 3806 instances which are not
solved by incompleteLAD, 1470 are solved by VF2 in less than 50 milliseconds. Therefore, after
running incompleteLAD, we run the VF2 solver for \SI{50}{\ms}. This solves easy instances without
the overhead of running algorithm selection and avoids potentially making incorrect solver choices.
After running incompleteLAD and VF2 for \SI{50}{\ms}, we are left with 2336 hard instances that we
consider for algorithm selection.

@@

LLAMA's pairwise regression approach with random forest regression gave the best performance. The
idea is very similar to the pairwise classification used by Xu et al.\ \cite{xu_satzilla_2008}. For
each pair of algorithms in our portfolio, we train a model that predicts the performance difference
between them. That is, if the first algorithm is better than the second, the difference is positive,
otherwise negative. The algorithm with the highest cumulative performance difference, i.e.\ the most
positive difference over all other algorithms, is chosen to be run.

@@

Table~\ref{tab:res} shows the performance of our algorithm selection approach, compared to two
baselines. The virtual best solver is the oracle predictor that, for each instance, chooses the best
solver from our portfolio. This is the upper bound of what an algorithm selection approach can
achieve. The single best solver is the one solver from the portfolio that has the overall best
performance across the entire set of instances, at the CPU time limit of \SI{10e8}{\ms},
i.e.\ Glasgow(2). We consider it a lower bound on performance.
We are able to solve 34 more instances than the single best solver within the timeout, with only an
additional 13 to the virtual best. In terms of average performance, we are able to close almost
\SI{70}{\percent} of the gap between the single best and the virtual best solver.

\begin{table}[ht]
    \centering\setlength{\tabcolsep}{1em}
\begin{tabular}{lrrr}
  \toprule
model & mean MCP & solved instances & mean performance\\
  \midrule
virtual best & 0 & 2219 & 5822809\\
single best & 1965158 & 2172 & 7787967\\
LLAMA & 610835 & 2206 & 6435301\\
   \bottomrule \\
\end{tabular}
\caption{Algorithm selection performance on the reduced set of 2336
instances. MCP is the misclassification penalty; that is, the additional time
required to solve a set of instances because of choosing solvers that perform
worse than the best.}\label{tab:res} \end{table}

Table~\ref{tab:res} shows the performance of the selection model. The performance of the entire
system we propose here is shown in Table~\ref{tab:resfull}. The additional instances are solved
either by incompleteLAD or VF2 presolving, as explained above. Our system is able to close more than
\SI{50}{\percent} of the gap between single and virtual best.
